\vspace{1.0cm} \Large \bf Integrating influenza antigenic dynamics with molecular evolution

### 2013-12-10 correspondence/reresponse.tex
Fix #34 - Ne*tau description.
Fix #35 - drift vs sampling count correlation.

@@

>>>>

\textbf{We've revised the manuscript to clarify the definition and interpretation of both $N_e$ and $\tau$.}

@@

\textbf{We've expanded out this sentence to a paragraph following the paragraph on the year-to-year drift vs incidence correlation.  We give rationale for the test, explain the bootstrap procedure, give $p$-values for both combined and separate analyses across lineages and provide a figure showing a scatterplot of the data.}

### 2013-12-10 flux.tex
Fix #35 - drift vs sampling count correlation.

We sought to summarize year-to-year patterns of antigenic drift by calculating the difference in mean location between consecutive years (Figure~\ref{jumps}A).
We estimate year-to-year antigenic drift for years 1992 to 2011 by calculating the average location along dimension 1 of phylogenetic lineages present in the tree at year $i$ and comparing this location to the average location of phylogenetic lineages present in the tree at year $i-1$.
There may often be large discontinuities in virus locations across the population; our use of difference in mean location is meant to capture both the distance between antigenic clusters and also the change in cluster frequency over consecutive years.
We observe greater heterogeneity in year-to-year antigenic drift in type A than in type B lineages (Figure~\ref{jumps}B), with standard deviation of year-to-year antigenic drift equal to 0.97 units in A/H3N2, 0.66 units in A/H1N1, 0.46 units in B/Vic and 0.26 units in B/Yam.
This analysis classifies drift only to the level of consecutive years; some coarse-graining of the timings of transition events will necessarily occur.
We tested to see if there could be an artifactual effect on the measurement of year-to-year drift based on the number of viral isolates collected by correlating drift between years $i$ and $i-1$ against the ratio of the number of isolates from year $i$ to the number of isolates from year $i-1$.
We found no significant correlation (Pearson's $r = -0.01$, $p = 0.411$), suggesting our results to be unbiased with regard to sample count.

@@

>>>>

We sought to summarize year-to-year patterns of antigenic drift by calculating the difference in mean location between consecutive years (Figure~\ref{jumps}A).
We estimate year-to-year antigenic drift for years 1992 to 2011 by calculating the average location along dimension 1 of phylogenetic lineages present in the tree at year $i$ and comparing this location to the average location of phylogenetic lineages present in the tree at year $i-1$.
There may often be large discontinuities in virus locations across the population; our use of difference in mean location is meant to capture both the distance between antigenic clusters and also the change in cluster frequency over consecutive years.
We observe greater heterogeneity in year-to-year antigenic drift in type A than in type B lineages (Figure~\ref{jumps}B), with standard deviation of year-to-year antigenic drift equal to 0.97 units in A/H3N2, 0.66 units in A/H1N1, 0.46 units in B/Vic and 0.26 units in B/Yam.
This analysis classifies drift only to the level of consecutive years; some coarse-graining of the timings of transition events will necessarily occur.

@@

However, it is possible that if sampling count influences cartographic estimates then a spurious correlation could arise in which years with greater incidence have higher sample counts and artifactually high estimates of drift.
We controlled for this possibility by testing to see if viral isolate count influences estimates of year-to-year drift by correlating drift between years $i$ and $i-1$ against the ratio of the number of isolates from year $i$ to the number of isolates from year $i-1$.
We found no significant correlation when combining data across lineages (Figure~\ref{counts_corr}, Pearson's $r = -0.01$).
We testing significance following the same bootstrap procedure we used to assess the correlation between drift and incidence, finding a $p$-value of 0.514.
Separating lineages gave $p$-values of 0.717, 0.246, 0.337, 0.504 for A/H3N2, A/H1N1, B/Vic and B/Yam, respectively.
These findings suggest our results to be unbiased with regard to sample count.

%%% counts_corr %%%
\begin{figure}[h]
	\centering		
	\includegraphics[width=0.4\textwidth]{figures/counts_corr}
	\caption{\textbf{Relationship between antigenic drift and sample counts for years 1998 to 2011 in influenza A/H3N2, A/H1N1, B/Vic and B/Yam.}
	Antigenic drift from year $i-1$ to year $i$ is compared to the ratio of sample counts in year $i$ to counts in year $i-1$.
	Only comparisons which had one or more samples in years $i-1$ and $i$ were retained, leaving 11 A/H3N2, 7 A/H1N1, 9 B/Vic and 10 B/Yam comparisons.
	Points are colored according to influenza lineage based on the color scheme in Figure \ref{corr}.
	} 
	\label{counts_corr} 
\end{figure}
